---
title: "Elin"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Can we do BLAST and annotation in R?

```{r
readFastq("MC2_1.fq")
```


## Find the coding ORFs of samples
Installing the ORFik package to find the ORFs of the coding regions to be able
to find proteins and annotate these.

```{r}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("ORFik")
```

Want to install the ShortRead package.. But it does not seem to work...
```{r
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("ShortRead")
```


```{bash}
R CMD INSTALL ShortRead_1.52.0.tar.gz
```
I also tried to install the library (BiocManager) on the path /home/student9/R/x86_64-pc-linux-gnu-library/4.1 so that i could maybe install 
the ShortString package, but to no use...

testing testing...
```{r}
source("http://www.bioconductor.org/biocLite.R")
biocLite(c("ShortRead"))
```

```{r}
install.packages(ShortReads)
```

## Finding the ORFs

Sandra installed it for all of us, wohoo!

```{r}
library(ORFik)
library(ShortRead)
library(GenomicFeatures)
library(stringr)

#?findORFs
```


```{r
MC2_1 <- readFastq("test/MC2_1.fastq")
```
Remade the paired fastq files into fasta files to be able to read them in R (and also find the ORFs)
```{bash}
conda activate filter #here I have the fastx_toolbox
cd home/student6/data-merged/
#ex:
fastq_to_fasta -i MC5_paired.fastq -o MC5_paired.fasta
# do with all files. 
```

Genome
```{r}
reference_medulis <- readDNAStringSet("/home/student6/data/GCA_019925275.1_PEIMed_genomic.fna")
```


Samples
```{r}
MC5_paired <- readDNAStringSet("/home/student6/data-merged/MC5_paired.fasta")
MI5_paired <- readDNAStringSet("/home/student6/data-merged/MI5_paired.fasta")
MO1_paired <- readDNAStringSet("/home/student6/data-merged/MO1_paired.fasta")
S221_paired <- readDNAStringSet("/home/student6/data-merged/S221_paired.fasta")
S323_paired <- readDNAStringSet("/home/student6/data-merged/S323_paired.fasta") 
```


```{r}
ORFS_MC5_paired <- findORFs(MC5_paired)
ORFS_MI5_paired <- findORFs(MI5_paired)
ORFS_MO1_paired <- findORFs(MO1_paired)
ORFS_S221_paired <- findORFs(S221_paired)
ORFS_S323_paired <- findORFs(S323_paired)
```

Finding the transcripts

```{r}
ORFseq_MC5_paired <- extractTranscriptSeqs(MC5_paired, ORFS_MC5_paired)
```

testingtesting
```{r}
test <- rbind(c(1,5,5), c(2,8,5), c(3,7,4))
testseq <- c("ABCDEFGHIJKLMN")

i=1
for(i in 1:3){
  a <- test[i,]
 print(str_sub(testseq,a[1],a[2]))
}


```

Need to make the fuincking IRange object into a vector.......
```{r}
start_MC5 <- start(ORFS_MC5_paired)
end_MC5 <- end(ORFS_MC5_paired)
range_MC5 <- cbind(start_MC5, end_MC5)

blopp <- as.matrix(ORFS_MC5_paired)
```


```{r}
for (i in 1:length(range_MC5)) {
  b <- range_MC5[i,]
  ORFseq_MC5_paired <- str_sub(MC5_paired, b[1], b[2])
}
```
**I think she got it! By George she's got it!**
```{bash}
# conda create -n orfipy -c bioconda orfipy
# conda activate orfipy
orfipy MO1_paired.fasta --dna ORFS_MO1.fa --min 201 --max 30590
# since we had to do this step in another way, we set the min and max values of
#the ORFs to the values that they found 
#then I put all the ORF_id.fa files in ORFS directory
```

Now I want to translate the DNA ORFs into amino acid orfs --> installing emboss so I can use transeq
```{bash}
conda create -n emboss
conda install -c bioconda emboss
conda activate emboss

transeq ORFS_MC5.fa ORFS_MC5.aa #making the aa files

#moving all to a new directory
mkdir gene_predicition
mv ORFS/*.aa gene_prediction/ #moving all files with ending .aa

#in the gene_prediction directory
conda activate gene_prediction
wget https://ftp.uniprot.org/pub/databases/uniprot/current_release/knowledgebase/complete/uniprot_sprot.fasta.gz
gunzip uniprot_sprot.fasta.gz

makeblastdb -in uniprot_sprot.fasta -dbtype prot -out uniprot_database

#then i run the blast p for each file.. will take a looong time. 
blastp -query ORFS_MC5.aa -db uniprot_database -outfmt 7 -out blastp_MC5

```


I now want to test if I get as many ORFs as they do?
```{bash}

```




